---
title: "Line distribution in movies"
subtitle: 'Gender issues in Line distribution in movies'
author: "José Benardi de Souza Nunes"
date: 29/06/2018
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
---

<br>

# Introduction

<br>

> Data analysis with non-hierarchical clustering (algorithm k-means) on data about line distribution in movies. The analysis was made on the dataset **Polygraph's Film Dialogue**. Information about the dataset and how it was generated can be found in its [original repository](https://github.com/matthewfdaniels/scripts).

<br>

***

<br>

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}

library(here)
library(broom)
library(vegan)
library(mclust)
library(plotly)
library(NbClust)
library(lattice)
library(cluster)
library(tidyverse)
library(ggfortify)

theme_set(theme_bw())
```

# Data Overview

```{r, warning=FALSE}
readr::read_csv(here("data/character_list5.csv"),
                      progress = FALSE,
                      col_types = cols(
                                    script_id = col_integer(),
                                    imdb_character_name = col_character(),
                                    words = col_integer(),
                                    gender = col_character(),
                                    age = col_character()
                                    )) %>%
  mutate(age = as.numeric(age)) -> characters_list

readr::read_csv(here("data/meta_data7.csv"),
                      progress = FALSE,
         col_types = cols(
                        script_id = col_integer(),
                        imdb_id = col_character(),
                        title = col_character(),
                        year = col_integer(),
                        gross = col_integer(),
                        lines_data = col_character()
                        )) %>%
  mutate(title = iconv(title,"latin1", "UTF-8")) -> meta_data
```

<br>

#### Combining original data

```{r}
left_join(characters_list, 
          meta_data, 
          by=c("script_id")) %>%
  group_by(title, year) %>%
  drop_na(gross) %>%
  ungroup() -> scripts_data

scripts_data %>%
  glimpse()
```

```{r}
scripts_data %>%
  mutate(fem_words = ifelse(gender == "f",words,0),
         man_words = ifelse(gender == "m",words,0)) %>%
  group_by(title, year) %>%
  mutate(total_fem_words = sum(fem_words),
         total_man_words = sum(man_words)) %>%
  filter(total_fem_words !=  0) %>%
  filter(total_man_words !=  0) %>%
    mutate(f_m_ratio = sum(gender == "f")/sum(gender == "m"),
           f_m_wordratio = total_fem_words/total_man_words) %>%
  ungroup()  -> scripts_data

scripts_data %>%
  select(title,
         year,
         f_m_ratio,
         f_m_wordratio) %>%
  sample_n(10)
```

## Exploring data

### Proportion of female and male dialogue

```{r}
scripts_data %>%
  group_by(title,year) %>%
  slice(1) %>%
  unique() %>%
  ggplot(aes(x=f_m_wordratio,
             y=(..count..)/sum(..count..))) +
  geom_histogram(binwidth = 0.1,
                 boundary = 0,
                 fill = "grey",
                 color = "black") +
  labs(y="Relative Frequency",
       x="female/male  wordratio")
```

* Em alguns raríssimos exemplos há muito mais dialógo feminino que feminino. 

```{r}
scripts_data %>%
  group_by(title,year) %>%
  slice(1) %>%
  unique() %>%
  filter(f_m_wordratio < 10) %>%
  ggplot(aes(x=f_m_wordratio,
             y=(..count..)/sum(..count..))) +
  geom_histogram(binwidth = 0.1,
                 fill = "grey",
                 color = "black") +
  labs(y="Relative Frequency",
       x="female/male  wordratio")
```

* Once the more unusual cases have been filtered we can see a strong dominance of masculine dialogue over feminine dialogue in the movies.

```{r}
scripts_data %>%
  group_by(title,year) %>%
  slice(1) %>%
  unique() %>%
  ggplot(aes(x="", 
             y=f_m_wordratio)) +
  geom_violin(fill="grey",
               width=0.5) +
  labs(y="female/male  wordratio") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

* It becomes even more apparent:
    * The existence of some few cases of complete dominance of feminine dialogue
    * The overall dominance of masculine dialogue over feminine dialogue.

### Proportion of female and male characters

```{r}
scripts_data %>%
  group_by(title,year) %>%
  slice(1) %>%
  unique() %>%
  ggplot(aes(x=f_m_ratio,
             y=(..count..)/sum(..count..))) +
  geom_histogram(binwidth = 0.1,
                 boundary = 0,
                 fill = "grey",
                 color = "black") +
  scale_x_continuous(breaks = seq(0,10,0.5)) +
    labs(y="Relative Frequency",
       x="(female chars / male chars) ratio")
```

* The predominance of male characters is clear

```{r}
scripts_data %>%
  group_by(title,year) %>%
  slice(1) %>%
  unique() %>%
  ggplot(aes(x="", 
             y=f_m_ratio)) +
  geom_violin(fill="grey",
               width=0.5) +
  labs(y="(female chars / male chars) ratio") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

* Besides the strong predominance of male characters we can see the existence in some instances however unfrequent of a overwhelming female presence (e.g 10 times more female than male characters).

### Year of release

```{r}
scripts_data %>%
  group_by(title,year) %>%
  slice(1) %>%
  unique() %>%
  ggplot(aes(x=year)) +
  geom_bar(fill = "grey",
           color = "black") +
  labs(y="Absolute Frequency",
       x="Year of release")
```

* Most of the movies are recent, almost all of them released not before the 1990s.

```{r}
scripts_data %>%
  group_by(title,year) %>%
  slice(1) %>%
  unique() %>%
  ggplot(aes(x="", 
             y=year)) +
  geom_violin(fill="grey",
               width=0.5) +
  labs(y="Year of release") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

* Ainda é possível ver uma presença relevante de filmes do começo dos anos 1980.
* Existem alguns filmes anteriores aos próprio anos 1950.

### Movie revenue

```{r}
scripts_data %>%
  group_by(title,year) %>%
  slice(1) %>%
  unique() %>%
  ggplot(aes(x=gross,
             y=(..count..)/sum(..count..))) +
  geom_histogram(binwidth = 50,
                 boundary = 0,
                 fill = "grey",
                 color = "black") +
  labs(y="Relative Frequency", x="Gross")
```

* Small or reasonable revenue for most movies.
* Some few movies had a overwhelming revenue.

```{r}
scripts_data %>%
  group_by(title,year) %>%   
  slice(1) %>%
  unique() %>%
  ggplot(aes(x="", 
             y=gross)) +
  geom_violin(fill="grey",
               width=0.5) +
  labs(y="Gross") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```

* Similar results to respective histogram.

## Putting data on appropriate scale.

```{r}
scripts_data %>%
  group_by(title) %>%
  slice(1) %>%
  unique() %>%
  ungroup() %>%
  select(title,
         gross,
         f_m_ratio,
         f_m_wordratio) -> data

select(data, -title) %>%
mutate_all(funs(scale)) -> scaled_data

scaled_data %>% 
  sample_n(10)
```

<br>

***

<br>

#  Optimal K for K-means

<br>

## GAP Statistic

The GAP Statistic compares the grouping results with each available k in a dataset where there isn't grouping structure. 

```{r}
plot_clusgap = function(clusgap, title="Gap Statistic calculation results"){
    require("ggplot2")
    gstab = data.frame(clusgap$Tab, k=1:nrow(clusgap$Tab))
    p = ggplot(gstab, aes(k, gap)) + geom_line() + geom_point(size=5)
    p = p + geom_errorbar(aes(ymax=gap+SE.sim, ymin=gap-SE.sim), width = .2)
    p = p + ggtitle(title)
    return(p)
}
```

```{r}
gaps <- scaled_data %>% 
    clusGap(FUN = kmeans,
            nstart = 20,
            K.max = 8,
            B = 200,
            iter.max=30)
```

```{r}
plot_clusgap(gaps)
```

* Both 3 groups (K = 3) and 6 groups (K= 6) seem appropriate, but as 6 is preceded by a downward trend 3 would be the choice.

## Elbow Method

```{r}
set.seed(123)
# Compute and plot wss for k = 2 to k = 15.
k.max <- 15

wss <- sapply(1:k.max, 
              function(k){kmeans(scaled_data, k, nstart=50,iter.max = 15 )$tot.withinss})
plot(1:k.max, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")
```

* According to the Elbow method 3 seems a good choice as the drop from 3 to 4 seems substantial.

## Bayesian Information Criterion

```{r echo=FALSE, message=FALSE}
d_clust <- Mclust(as.matrix(scaled_data), G=1:15,
                  modelNames = mclust.options("emModelNames"))

plot(d_clust$BIC)
```

* Visually K= 2 e K = 3 both represent the most meaningful gain in terms of BIC (Bayesian Information Criterion) 

## Hubert Index and D Index

```{r}
nb <- NbClust(scaled_data, diss=NULL, distance = "euclidean", 
              min.nc=2, max.nc=5, method = "kmeans", 
              index = "all", alphaBeale = 0.1)
hist(nb$Best.nc[1,], breaks = max(na.omit(nb$Best.nc[1,])))
```

* Hubert Index and D Index suggest K = 4 as the best solution

<br>

***

<br>

# K-Means 

<br>

## Agrupamento

```{r}
n_clusters = 3

scaled_data %>%
    kmeans(n_clusters, iter.max = 100, nstart = 20) -> km

p <- autoplot(km, data=scaled_data, frame = TRUE)  

ggplotly(p)

```

* É possível ver os 3 grupos nitidamente distintos, por meio de um zoom percebe-se que embora o grupo 1 e o grupo 3 estejam próximos o overlap é basicamente inexistente.

```{r, warning=FALSE}
row.names(scaled_data) <- data$title

toclust <- scaled_data %>% 
    rownames_to_column(var = "title") 

km = toclust %>% 
    select(-title) %>% 
    kmeans(centers = n_clusters, iter.max = 100, nstart = 20)

km %>% 
    augment(toclust) %>% 
    gather(key = "variável", value = "valor", -title, -.cluster) %>% 
    ggplot(aes(x = `variável`, y = valor, group = title, colour = .cluster)) + 
    geom_point(alpha = 0.2) + 
    geom_line(alpha = .5) + 
    facet_wrap(~ .cluster) +
    coord_flip()

```

<br>

***

$\color{red}{\text{Grupo 1}}$ - **We Can Do It!**

  * Menor Faturamento 
  * Mais dialógo para as mulheres 
  * Maior taxa de personagens femininos
    
<br>

O $\color{red}{\text{Grupo 1}}$ - **We Can Do It!** é o grupo de filmes de maior representação feminina, quer seja em proporção de personagens femininos como em proporção de dialógos dedicados a personagens femininos. Existe porém uma característica negativa que acompanha este mesmo grupo, pois este é também o grupo das menores taxas de faturamento. Isso sugere uma infeliz associação negativa entre a representação feminina em filmes e o faturamento destes. 

<br>


```
O nome do grupo se refere ao famoso cartaz de J. Howard Miller de 1943 incentivado as mulheres a participar no esforço de guerra nas fábricas. 
```

<br>

***

$\color{green}{\text{Grupo 2}}$ - **It's A Man's Man's Man's World**   

* Maior faturamento entre todos
* Menor taxa de dialógo para as mulheres
* Menor taxa de personagens femininos
    
<br>

O $\color{green}{\text{Grupo 2}}$ - **It's A Man's Man's Man's World** é o grupo de filmes de menor representação feminina, quer seja em proporção de personagens femininos como em proporção de dialógos dedicados a personagens femininos. Existe porém uma característica negativa que acompanha este mesmo grupo, pois este é também o grupo de maiores taxas de faturamento. Isso sugere uma infeliz associação positiva entre ausência de  representação feminina em filmes e o faturamento destes. 

<br>

```
O nome do grupo se refere à música de James Brown, a qual foi escrita por sua então namorada Betty Jean Newsome como um comentário sobre a relação entre os sexos.
```

<br>

***

$\color{blue}{\text{Grupo 3}}$ - **Em cima do muro**

  * Filmes medianos em termos de proporção de personagens femininos, proporção de dialógos dedicados a personagens femininos e faturamento.
    
<br>

```
O nome do grupo se refere à expressão que significa não tomar partido.
```

<br>

## Qualidade da clusterização / Silhueta

```{r}
dists = scaled_data %>% 
  dist()

scaled_data %>%
    kmeans(3, iter.max = 100, nstart = 20) -> km


silhouette(km$cluster, dists) %>%
   plot(col = RColorBrewer::brewer.pal(4, "Set2"),border=NA)
```

<br>

* O valor de 0.6 da silhueta significa que a nossa clusterização foi razoável. ヾ(⌐■_■)ノ♪

